---
title: "Conversion Hotspots"
author: "Alex Zvoleff"
date: "2024-06-08"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(tictoc)
library(rnaturalearth)
library(rnaturalearthdata)
library(cowplot)
library(lwgeom)

world <- ne_countries(scale = "medium", returnclass = "sf")
crs_wintri <- "+proj=wintri +datum=WGS84 +no_defs +over"
world_wintri <- st_transform_proj(world, crs = crs_wintri)
grat_wintri <- 
  st_graticule(lat = c(-89.9, seq(-60, 60, 30), 89.9),
               lon = c(-179.9, seq(-135, 135, 45), 179.9)) %>%
  st_transform_proj(crs = crs_wintri)
```

```{r read_data, cache=TRUE}
hotspots <- read_csv("conversionHotspots.csv")

ecoregions_raw <- st_read("ecoregions/Ecoregions2017.shp")
ecoregions_raw <- st_make_valid(ecoregions_raw)
ecoregions <- st_simplify(ecoregions_raw, dTolerance = 1000)
ecoregions <- left_join(ecoregions, hotspots, by=join_by(ECO_NAME==ecoregion))

# Check that there are no NAs (i.e. that the join worked correctly)
table(is.na(ecoregions$allVegConversion00_10))
table(is.na(ecoregions$allVegConversion10_20))

```

## Setup indicators of top ecoregions in terms of conversion

Look at top 10% of ecoregions in terms of recently converted area (from 2010 -
2020), and also in terms of total conversion since 2000. Also add an indicator
of ecoregions where conversion has accelerated (i.e. higher rates of conversion
in 2010-2020 version 2000-2010).

```{r calculate_top}
ecoregions %>%
    filter(!is.na(allVegArea00)) %>%
    mutate(accelerating=ifelse(allVegConversion10_20 > (allVegConversion00_10*1.1),
                               'Accelerating', 'Decelerating'),
           topByAreaRecent=percent_rank(allVegConversion10_20)>.9,
           topByPctRecent=percent_rank(allVegConversion10_20/(allVegArea00 - allVegConversion00_10))>.9,
           topByAreaOverall=percent_rank(totalAllVegConversion00_20)>.9,
           topByPctOverall=percent_rank(totalAllVegConversion00_20/allVegArea00)>.9) -> ecoregions
```


### Plot some basic stats

``` {r basic_stats, echo=FALSE}
ecoregions %>%
    ggplot() +
    geom_histogram(aes(allVegArea00))

ecoregions %>%
    ggplot() +
    geom_histogram(aes(allVegConversion10_20))

ecoregions %>%
    ggplot() +
    geom_histogram(aes(allVegConversion10_20/(allVegArea00 - allVegConversion00_10)))

ecoregions %>%
    ggplot() +
    geom_histogram(aes(totalAllVegConversion00_20))

ecoregions %>%
    ggplot() +
    geom_histogram(aes(totalAllVegConversion00_20/allVegArea00))

ecoregions %>%
    ggplot(aes(allVegConversion10_20)) +
    stat_ecdf(geom = "step")

ecoregions %>%
    ggplot(aes(allVegConversion10_20/allVegArea00)) +
    stat_ecdf(geom = "step")

```

### What percentage of total conversion does each method capture?

``` {r percent_of_total_conv}
# Recent conversion
sum(ecoregions$allVegConversion10_20[ecoregions$topByPctRecent], na.rm=TRUE) / sum(ecoregions$allVegConversion10_20) * 100
sum(ecoregions$allVegConversion10_20[ecoregions$topByAreaRecent], na.rm=TRUE) / sum(ecoregions$allVegConversion10_20) * 100

# Overall conversion
sum(ecoregions$totalAllVegConversion00_20[ecoregions$topByPctOverall], na.rm=TRUE) / sum(ecoregions$totalAllVegConversion00_20) * 100
sum(ecoregions$totalAllVegConversion00_20[ecoregions$topByAreaOverall], na.rm=TRUE) / sum(ecoregions$totalAllVegConversion00_20) * 100
```
### How many ecoregions show up in multiple methods
``` {r sensitivity}
ecoregions$nTop <- ecoregions$topByPctRecent + ecoregions$topByAreaRecent + ecoregions$topByPctOverall + ecoregions$topByAreaOverall
table(ecoregions$nTop)
```


## And what do maps look like

### By area, last 10 years
``` {r map_top_area_recent}
ggplot() +
  geom_sf(data=world_wintri) +
  geom_sf(data=filter(ecoregions, topByAreaRecent), aes(fill=ECO_NAME)) +
  geom_sf(data=grat_wintri, color = "gray90", size = 0.05/.pt) + 
  coord_sf(datum=st_crs("+proj=wintri +datum=WGS84 +no_defs +over")) +
  theme_map() +
  theme(legend.position="none")
```

### By percent converted, last 10 years
``` {r map_top_percent_recent}
ggplot() +
  geom_sf(data=world_wintri) +
  geom_sf(data=filter(ecoregions, topByPctRecent), aes(fill=ECO_NAME)) +
  geom_sf(data=grat_wintri, color = "gray90", size = 0.05/.pt) + 
  coord_sf(datum=st_crs("+proj=wintri +datum=WGS84 +no_defs +over")) +
  theme_map() +
  theme(legend.position="none")
```

### By area converted, last 20 years

``` {r map_top_area_overall}
ggplot() +
  geom_sf(data=world_wintri) +
  geom_sf(data=filter(ecoregions, topByAreaOverall), aes(fill=ECO_NAME)) +
  geom_sf(data=grat_wintri, color = "gray90", size = 0.05/.pt) + 
  coord_sf(datum=st_crs("+proj=wintri +datum=WGS84 +no_defs +over")) +
  theme_map() +
  theme(legend.position="none")
```

### By percent converted, last 20 years

``` {r map_top_percent_overall}
ggplot() +
  geom_sf(data=world_wintri) +
  geom_sf(data=filter(ecoregions, topByPctOverall), aes(fill=ECO_NAME)) +
  geom_sf(data=grat_wintri, color = "gray90", size = 0.05/.pt) + 
  coord_sf(datum=st_crs("+proj=wintri +datum=WGS84 +no_defs +over")) +
  theme_map() +
  theme(legend.position="none")
```

## The 21 ecoregions that show up across methods

``` {r map_top_across_methods}
ecoregions %>%
  filter(nTop==4) %>%
  arrange(ECO_NAME) %>%
  st_drop_geometry() %>%
  select(ECO_NAME)

ggplot() +
  geom_sf(data=world_wintri) +
  geom_sf(data=filter(ecoregions, nTop==4), aes(fill=ECO_NAME)) +
  geom_sf(data=grat_wintri, color = "gray90", size = 0.05/.pt) + 
  coord_sf(datum=st_crs("+proj=wintri +datum=WGS84 +no_defs +over")) +
  theme_map() +
  theme(legend.position="none")
```

``` {r intersect_area, cache=TRUE}
gadm <- st_read("C:/Users/azvoleff/OneDrive - Conservation International Foundation/Data/gadm_410-levels.gpkg", "ADM_1")
gadm_0 <- st_read("C:/Users/azvoleff/OneDrive - Conservation International Foundation/Data/gadm_410-levels.gpkg", "ADM_0")
gadm <- st_make_valid(gadm)

ecoregions$hotspot <- ecoregions$nTop==4

tic()
gadm %>%
  st_filter(y=filter(ecoregions, hotspot), .predicate = st_intersects) %>%
  mutate(admin_area = st_area(.)) -> hotspot_admin
gadm_intersection <- st_intersection(hotspot_admin, filter(ecoregions, hotspot))
toc()

full_join(
  hotspot_admin,
  gadm_intersection %>%
    mutate(intersect_area = st_area(.)) %>%
    select(GID_1, intersect_area) %>%
    st_drop_geometry() %>%
    group_by(GID_1) %>%
    summarise(intersect_area=sum(intersect_area))
  ) %>%
  mutate(intersect_frac=as.numeric(intersect_area / admin_area)) -> intersection_frac
```

## Jurisdictions overlapping the 21 ecoregions that show up across methods

### Jurisdictions by percent overlap
``` {r plot_intersect_jurisdictions}
ggplot() +
  geom_sf(data=world_wintri) +
  geom_sf(data=intersection_frac, aes(fill=intersect_frac > .5)) +
  geom_sf(data=grat_wintri, color = "gray90", size = 0.05/.pt) + 
  coord_sf(datum=st_crs("+proj=wintri +datum=WGS84 +no_defs +over")) +
  theme_map()
```

### Jurisdictions with at least 50% overlap overlaid on hotspot ecoregions
``` {r plot_intersect_jurisdictions_on_hotspots}
ggplot() +
  geom_sf(data=world_wintri) +
  geom_sf(data=filter(ecoregions, hotspot), fill='cyan') +
  geom_sf(data=filter(intersection_frac, intersect_frac > .5), fill='blue') +
  geom_sf(data=grat_wintri, color = "gray90", size = 0.05/.pt) + 
  coord_sf(datum=st_crs("+proj=wintri +datum=WGS84 +no_defs +over")) +
  theme_map()

```